A new simplified manual tour, with application to exploring classification boundaries in high dimensions

### 2022-07-07 paper/paper.tex
compiled tex file

>>>>

We use the \texttt{classifly} package \citep{classifly} to generate
predictions across the 4D space spanned by the data, with two
classification models (linear discriminant analysis and random forest).

The first projection we will start from is looking at bill depth vs
flipper length (x2 vs x3). In a projection of the data we see that the
Gentoo penguins are separate from the other two species which are on top
of each other. If we instead look at a slice through the center of the
distribution, we see that shifting the center point to large values of
x1 (bill length), e.g.~x1=2, results in Adelie penguins dropping out of
the slice, while a slice with low values (x1=-2) in the center point
only contains Adelie penguin observations.

We can next explore how this is mapped in the two classification models.
Both random forest and LDA can separate out the region for Gentoo
penguins in a thin slice through the center (h = 0.5), while Adelie and
Chinstrap predictions overlap (blue and yellow, an effect of the slice
thickness). We also notice the different boundaries: while LDA produces
linear boundaries by construction, random forest insead produces
``boxes'' from cutting up the data in the tree structure.

\begin{figure*}[ht]
\centerline{\includegraphics[width=0.4\textwidth]{figures/view1_lda.png}
\includegraphics[width=0.4\textwidth]{figures/view1_rf.png}}
\caption{Model predictions in a centered slice defined by the projection onto bill depth vs flipper length, from an LDA and RF model. The colors are the same as used in the SPLOM, we see that Gentoo penugins are clearly separated, while the other two species are overlapping, and we can see the distinugishing features of the two models in terms of the different boundaries.}
\label{view1}
\end{figure*}

We can then start shifting the slice along x1 as we did when looking at
the data, again moving to \(\pm 1\) will separate the Adelie from the
Chinstrap penguins, now in terms of the model predictions. (Could
include screenshot here as well.)

Thus, stepping through the different slices shows the relation between
the species and the first variable (x1, bill length). Instead of moving
the center point we could also keep the slice fixed as passing through
the center of the distribution, and use the manual controls to rotate x1
into the projection. Here we started from the projected view of the data
to identify a combination that can separate all three species, and then
use the projection matrix as input to the slice displays for the model.
Interestingly, the two models agree on how to classify the ``empty''
regions (where there are no data points) for this example. (this is
applying A3 as the starting projection).

Finally we use the manual tour on the random forest predictions first,
to identify a centered slice with an interesting classification
boundary. Here a combination of all four predictors is found to result
in an interesting slice, where the predictions for the Gentoo penguins
gets split into two regions of the slice, indicating the non-linearity
of the random forest model. While x4 (body mass) is already contributing
a lot in the projection, we can learn more by also moving the center
point along that direction. Moving to high values (x4 = 2) results in a
much larger part of the slice being predicted as Gentoo penguins, while
now the yellow region (Chinstrap penguins) gets split. Moving in the
opposite direction (x4 = -1) instead shows primarily Chinstrap penguins
in the predictions, with Gentoo only appearing along the edge.

Looking at the data, and shifting a fat slice (h = 1.5) as specified
above (x4 = -1, 0, 2) we see that slices through x4 = -1, 0 do not have
observations from the Gentoo species, while x4 = 2 contains the Gentoo
observations, but hardly any of the Chinstrap penguins.

The last step is comparing to the LDA predictions in those same slices.
With RF we could see the non-linear aspects in the boundaries, while LDA
can only produce linear boundaries, so these look very different. On the
other hand, the same dependence is observed when shifting the slice
along the x4 direction.
